---
title: 'Natural Intelligence Team - MSDS 6306: Case Study 01'
author: "Ruben Chavez, Michael Wolfe, Ayoade Dare"
date: "January 5, 2019"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<center> <h1>Brewery Market Analysis for ExploringNewBusiness Brewery</h1> </center>
<center> <h1>Initial Report Addendum</h1> </center>

Natural Intelligence Analysis Inc would like to thank ExploringNewBusiness Breweries for allowing us to provide your market analysis of current breweries in the United States. Per our meeting following the original report we have prepared the following addendum to answer your questions. The following report can be run at any time if the provided Beers and Brewery datasets are updated.


### Brewery by state

Due to your recent decision to leave your current state and move the propective location of your business we have expanded our analysis to the entire United States. A list of the number of breweries in each state provided an intial idea of the market saturation of breweries labeled as "local" in each state.


####Number of Breweries per State
```{r Brewery by State}
#Import required libraries
suppressMessages(library(RCurl))
suppressMessages(library(readr))
suppressMessages(library(knitr))
suppressMessages(library(ggplot2))
suppressMessages(library(kableExtra))
suppressMessages(library(dplyr))
#read Beer and Brewery data 
BeersURL <- getURL("https://raw.githubusercontent.com/mjwolfe91/DDS_401_TeamNI_Case_Study1/master/Beers.csv")
BrewURL <- getURL("https://raw.githubusercontent.com/mjwolfe91/DDS_401_TeamNI_Case_Study1/master/Breweries.csv")
Beers <- read_csv(BeersURL)
Breweries <- read_csv(BrewURL)
#calculate and display number of breweries per state
BreweryPerState <- data.frame(table(Breweries$State))
colnames(BreweryPerState) <- c("State", "Number of Breweries")
kable(BreweryPerState, format = "html", align = "c") %>% kable_styling(bootstrap_options = c("striped", "hover"))
```


Further analysis of brewery location analysis in terms of brewery per capita is available upon request.
 
 
```{r Combine Beer and Brewery data}
#Combine Beer and Brewery data
BeersAndBreweries <- merge(Beers, Breweries, by.x = "Beer_ID", by.y = "Brew_ID")
colnames(BeersAndBreweries) <- c("Beer_ID", "Beer_Name", "ABV", "IBU", "Brewery_id", "Style", "Ounces", "Brewery_Name", "City", "State")
```


###Combination of Datasets

The following table provides a sample of the first ten and last ten beers with their respective breweries. These tables allows you to do a quick spot check of the dataset used by the report. This is necessary in case your source data should change formats going forward. 


```{r Print First and Last ten rows of Merged Data Frame}
#print head and tail of merged data
kable(head(BeersAndBreweries, 10), caption = "First Ten Beers of Merged DataFrame", align = "c")  %>% kable_styling(bootstrap_options = c("striped", "hover"))
kable(tail(BeersAndBreweries, 10), caption = "Last Ten Beers of Merged DataFrame", align = "c")  %>% kable_styling(bootstrap_options = c("striped", "hover"))
```


###Missing Values

We have noticed that the available data does include a number of missing values. We have provided a table to allow for line of site of the count of missing data per column. We highly recommend your marketing team continue to add to your beer and brewery initial market research data sets until the table shows all 0's to allow for the provided report to be increasingly accurate.


```{r count of missing values}
#Create data frame of count of NA and format for display
CountOfNA <- data.frame(colSums(is.na(BeersAndBreweries)))
colnames(CountOfNA) <- c("Missing Values")
CountOfNA$ColumnName <- row.names(CountOfNA)
rownames(CountOfNA) <- c()
CountOfNA <- CountOfNA %>% select("ColumnName", "Missing Values")
#Display table of NA dataframe
kable(CountOfNA)  %>% kable_styling(bootstrap_options = c("striped", "hover"))
```


###Beer Characteristics by State -- ABV and IBU

We understood your concern about differing tastes in beer style regarding Alcohol by Volume (ABV) and International Bitterness Unit (IBU).A median value for ABV and IBU by state provides a quick reference point until a more in depth analysis can be completed. 

*If any state has no reported Beers a relevant message will be produced before the graph.*
```{r median ABV and IBU by State, fig.width=20}
#Create Median IBU and ABV dataframes
MedianIBU <- as.data.frame.table(tapply(BeersAndBreweries$IBU, BeersAndBreweries$State, median, na.rm=TRUE))
colnames(MedianIBU) <- c("State", "MedianIBU")
MedianABV <- as.data.frame.table(tapply(BeersAndBreweries$ABV, BeersAndBreweries$State, median, na.rm=TRUE))
colnames(MedianABV) <- c("State", "MedianABV")

#Create plots of Median IBU and ABV
ggplot(data = MedianIBU, aes(x = State, fill = MedianIBU, y = MedianIBU), xlab("State")) + geom_bar(stat = "identity") + ggtitle("Median IBU by State") + xlab("State") + ylab("Median IBU") + labs(fill = "State") + theme(plot.title = element_text(hjust = 0.5, size = 30, face = "bold")) + scale_color_discrete(drop=FALSE)
ggplot(data = MedianABV, aes(x = State, y = MedianABV, fill = MedianABV), xlab("State")) + geom_bar(stat = "identity") + ggtitle("Median ABV by State") + xlab("State") + ylab("Median ABV") + labs(fill = "State") + theme(plot.title = element_text(hjust = 0.5, size = 30, face = "bold")) + scale_color_discrete(drop=FALSE)
```


A snapshot of the current beer with the highest ABV and IBU will allow you to ensure that your "Jupiter IPA" beer can still be marketed as the "highest alcohol by volume IPA with a punch of hops to the face".


####Beer with Highest ABV
```{r largest ABV}
#Identify and display the Beer with the highest ABV
maxABV <- BeersAndBreweries[which.max(BeersAndBreweries$ABV),]
kable(maxABV)  %>% kable_styling(bootstrap_options = c("striped", "hover"))
```


####Beer with Highest IBV
```{r largest IBV}
#Identify and display the Beer with the highestIBU
maxIBU <- BeersAndBreweries[which.max(BeersAndBreweries$IBU),]
kable(maxIBU)  %>% kable_styling(bootstrap_options = c("striped", "hover"))
```


Jupiter IPA should also be compared to the current national ABV in general.


####National ABV Statistics
```{r summart stats for ABV}
#Create Dataframe version of sumamry statistics and format
ABVSummary <- data.frame(levels(factor(summary(BeersAndBreweries$ABV))))
ABVSummary <- as.data.frame(t(ABVSummary))
row.names(ABVSummary) <- c()
colnames(ABVSummary) <- c(labels(factor(summary(BeersAndBreweries$ABV))))
#display sumamry statistics of ABV column
kable(ABVSummary)  %>% kable_styling(bootstrap_options = c("striped", "hover"))
```


###Relationship of ABV to IBU

We understand your concern that further increase to the ABV of Jupiter IPA will negatively affect the beer's IBU. The data currently shows no need for such a concern. In fact the IBU seems to increase with ABV, albeit with a relatively loose correlation.

```{r scatterplot relationship between ABV and IBU}
ggplot(data = BeersAndBreweries, aes(ABV, IBU)) + geom_point(stat = "identity", color='blue') + geom_smooth(color='red',data = BeersAndBreweries, aes(x=ABV, y=IBU)) + ggtitle("ABV and IBU Regression Analysis") + labs(fill = "ABV and IBU Regression Analysis") + theme(plot.title = element_text(hjust = 0.5, size = 30, face = "bold")) + scale_color_discrete(drop=FALSE)
```

##Future Follow-Up

From the above analysis, there's no clear causal relationship between ABV and IBU, despite the loose correlation among the two. We recommend further research into the link between ABV and IBU to ensure continued success in ExploringNewBusiness Brewery's endeavors in this market.
